Identifying Software Performance Changes Across Variants and Versions

### 2020-08-16 paper.tex
paper changes

	Our main idea is as follows: We address the configuration complexity of this problem by selecting representative sample sets of configurations.
	Next, we uniformly sample a constant number of commits for each configuration and conduct respective performance measurements.
	Based on these measurement data, we assemble a prediction model.
	For each configuration, we estimate the likelihood of each commit being a change point, i.e. that performance of some configurations changes abruptly compared to the previous commit.
	Next, we leverage similarities in the performance histories of configurations that share common options.
	Since we sample the commits for each configuration independently, for each change point, we obtain many estimations using measurements of different commits.
	Overall, this allows us to obtain more accurate estimations of performance-changing commits with tractable effort.
	Based on a mapping from configurations to predicted change point probabilities for each configuration and commit, we derive the options responsible for each particular change point: configuration-dependent change points.
	In summary, we offer the following contributions:
	\begin{itemize}
		\item A novel technique to effectively identify shifts in performance of configurable software systems. We are able to pinpoint causative commits and affected configuration options with high accuracy and tractable effort.
		\item A feasibility demonstration of our approach by implementing an adaptive learning strategy to obtain accurate estimations with acceptable measurement cost.
		\item An evaluation using both synthetic and real-world performance data from three configurable software systems. Synthetic data lets us assess our model and approach conceptually and at scale, whereas we are able to assess practicality with real-world data.
		\item A companion website\footnote{\url{https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions} or archived at {\color{red}\url{SOFTWAR-HERITAGE}}} providing supplementary material including a reference implementation of our approach, performance measurement data, and additional visualizations.
	\end{itemize}

@@

	\begin{figure}
		\includegraphics[width=0.5\textwidth]{images/sampling_plan}
		\label{fig:sampling plan}
		\caption{Active sampling strategies for a DBMS with 4 configurations}
	\end{figure}

@@

	{\color{magenta}We illustrate the measurement effort in Figure~\ref{fig:sampling_all}, where each dot represents a pair of configuration and version.
		To reduce measurement effort, one could limit the measurements to either a few configurations (cf. Figure~\ref{fig:sampling_configurations}) or a few commits (cf. Figure~\ref{fig:sampling_versions}), but at the cost of information loss along the respective complementary dimension.
		This would leave us with either accurate performance-influence models for few versions, resulting in poor temporal granularity, or with accurate performance histories, but only for a few configurations.
		Both do not provide accurate estimations of changes in the performance influence.}

	To obtain accurate estimations with a limited budget of measurements, we require a different approach. 
	The latter describes, in fact, an iterative and adaptive sampling approach.
	That is, we use an initial, but sparse sample set to explore the problem space and then increase the level of granularity at promising regions and dimensions (history segments and individual features and interactions, in our case).
	Adaptive sampling techniques have been successfully applied to  obtain both performance-influence models~\cite{siegmundPredictingPerformanceAutomated2012,sarkarCostEfficientSamplingPerformance} and performance histories~\cite{muhlbauer_accurate_2019} before.
	However, it is unclear whether fewer measurements are sufficient to assess the performance-influence of configuration options and interactions with respect to version changes. This is what we aim to accomplish with our work.

	\begin{figure}
		\captionsetup{format=hang}
		\begin{subfigure}{0.49\linewidth}
			\includegraphics[width=\linewidth]{images/sampling_1.pdf}
			\caption{Many configurations, but few versions}
			\label{fig:sampling_configurations}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.49\linewidth}
			\includegraphics[width=\linewidth]{images/sampling_2.pdf}
			\caption{Many versions, but few configurations}
			\label{fig:sampling_versions}
		\end{subfigure}
		\vfill
		\begin{subfigure}{0.49\linewidth}
			\includegraphics[width=\linewidth]{images/sampling_3.pdf}
			\caption{Many versions and configurations\\}
			\label{fig:sampling_all}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.49\linewidth}
			\includegraphics[width=\linewidth]{images/sampling_4.pdf}
			\caption{Initial sample augmented with additional measurements}
			\label{fig:sampling_both}
		\end{subfigure}

@@

	%	\begin{figure}
	%\centering
	%\includegraphics[width=\linewidth]{images/complexity.pdf}
	%\caption{Complexity}
	%\label{fig:complexity}
	%\end{figure}

@@

	\subsubsection{Candidate Commits}
	For each configuration, we compute an approximation of the change point likelihood over commits (cf. Figure~\ref{fig:cpp_example}) to identify local maxima (peaks). We select such peaks under the condition that the peak change point likelihood is greater than a  threshold $t_\text{CPL}$, as defined in Equation~\ref{eq:t_cpl}. This threshold is the average change point likelihood over all commits plus a factor $N_\text{CPL}$ times its corresponding standard deviation over all commits. 
	\begin{equation}
	\begin{gathered}
	t_\text{CPL} = \frac{1}{\vert V\vert} + N_\text{CPL} \cdot \sqrt{\frac{1}{\vert V\vert} \cdot \sum_{v \in V} \Big(\frac{1}{\vert V\vert} - p'(v)\Big)^2}
	\label{eq:t_cpl}
	\end{gathered}
	\end{equation}
	The factor $N_\text{CPL}$ (default value: 3) allows us to filter peaks that do not stand out enough and might be false positives. The greater $N_\text{CPL}$ is, the stricter the filtering of peaks. That is, for each configuration in our learning set, we obtain a set of commits (configuration-specific candidate commits) that represent possible change points.
	To reduce variation among the obtained commits, we cluster the commits using kernel density estimation (KDE). The purpose of KDE in our setting is to estimate the probability mass function of whether a commit is a change point. The local maxima of this KDE, subsequently, represent our set of candidate commits. {\color{red}Note that if two distinct change points are almost coincident commit-wise, they can be mistakenly identified by our approach as one change point that, subsequently, can be associated with configuration options belonging to the two individual change points. }

@@

	\subsubsection{Acquiring Configurations}
	The exploration of new configurations is similar to the initial sample selection strategy (distance-based sampling) described in Section~\ref{sec:initialization}. We select a constant number of configurations (default value: 75) for exploration. The exploitation of new configurations requires additional constraints. We define the following constraints for exploiting the candidate solution's associations. {\color{blue}For each commit in the candidate solution, we perform exploitation in two directions. For the first direction, at least 80\,\% of  configuration options associated with a commit are required to be selected, whereas, for the second direction, at least 80\,\% configuration options not associated with a commit are required to be selected. This way, we limit variation among configuration options. The first direction aims at enabling additional configuration options to search for previously unseen associations (under-approximation). The second direction aims at the opposite: We limit variation to those configuration options already associated with a commit and therefore search for configuration options that are falsely associated (over-approximation). We use the limit of 80\,\% to allow a broader scope of valid configurations to be sampled by a constraint solver. For each new configuration, we assess performance for randomly selected commits.}

@@

	To mitigate this trade-off in our evaluation, we conduct two separate experiments, based on synthetic and real-world performance measurements. The first experiments uses synthesized performance data and allows for a controlled experimental setup to assess scalability, accuracy, and efficiency at low cost while simultaneously be able to simulate different scenarios by varying the number of change points and affected configurations in the synthesized data.
	{\color{red}The second experiment uses a batch of real-world performance measurements of three software systems as a non-exhaustive ground truth to explore whether our algorithm can be practically applied to real-world systems.}

@@

	As a workload for the file compression tools \textsc{lrzip} and \textsc{xz}, we compressed the Silesia corpus which contains over 200~MB of files of different types.  {\color{red} This workload was designed to compare different file compression algorithms and tools.} For the audio transcoder \textsc{oggenc}, we encoded a raw WAVE audio file of over 60~MB from the Wikimedia Commons collection.  The workloads for \textsc{xz} and \textsc{lrzip} have been studied in previous work~\cite{siegmund_views_2015,muhlbauer_accurate_2019}. For all three subject systems, we assess performance by reporting the execution time.

@@

	\subsubsection{Operationalization}\label{sec:rq2_operationalization}
	For the experiment, we apply our approach on the performance measurements obtained in our pre-study with a few adjustments.
	First, although our study provides a broad and representative sample set of configurations, we cannot arbitrarily sample valid configurations for a couple of reasons. {\color{red}In particular, these include the commits that do not build as well as hidden configuration options and constraints. While we collected configuration options and constraints thoroughly from documentation artifacts, we can not assure that our selection is complete. We discuss this limitation further in the threats to validity.}

	Instead, when acquiring new configurations (cf. Sec.~\ref{sec:acquisition}), we select configurations {\color{red} from our batch of measurements} that (1) have not already been used by our algorithm, and (2) are "closest" to the requested configuration (i.e., with the minimal Hamming distance). The rationale is that this allows for rapid repetitions with different initializations {\color{red}of our algorithm}. Since we employ multiple sampling strategies in our pre-study we are confident that our broad batch of measurements is representative. Second, we set the number of measurements per iteration to 300. This value is substantially smaller than in our controlled experiment setup, where we explore the scalability of our approach. Third, as an initial configuration sample set, we randomly sample 5 configurations from our pre-study set. 
	Last, we initialize our approach with a relative performance threshold $\theta_\tau$ (cf. Equation~\ref{eq:theta}) of ten percent. In the context of the reported relative variation among performance measurements on the machines used, we consider this a rather conservative threshold. 
	We repeat the experiments 10 times with different seeds to assess robustness of our approach and account for the randomness in exploring the configuration space.
	We assess practicality by reporting measurement effort, number of iterations in relation to the pre-study's results, which are based on a vastly greater measurement budget.

@@

	\section{Discussion}
	\paragraph{Efficiency} Among all experiments conducted with our synthetic setup, our approach yielded reasonable to high accuracy across most parameter settings. A subsequent analysis of parameter settings showed that our algorithm is not limited by the size of a problem space, but rather the number of measurements per iteration. Our algorithm found multiple change points across settings of up to 2,500 commits and and 2$^{64}$ configurations, resulting in a search space of about $4.6 \cdot 10^{22}$ possible measurements.
	In the context of not more than 5,000 measurements per iteration, the maximum number of measurements (after 30 iterations) added to the initial set of measurements is 150,000. For a setting with 2,500 commits, this represents around $1.4 \cdot 10^{-6}$\,\% and $3.2\cdot 10^{-16}$\,\% of possible measurements for settings with 32 and 64 options, respectively. The vast majority of cases terminated in early iterations. Hence, we are confident that our approach scales and can efficiently approximate change points for large software systems. We, we are able to reproduce this finding with real-world performance when identifying the temporal location of change points. {\color{red}For our controlled experiment setup, we have not tuned the algorithms’ parameters to reach optimal efficiency or accuracy. However, without tuning, we demonstrated the feasibility and efficiency of our approach. Project-specific fine tuning of several parameters is a possible extension to our approach.}

	\paragraph{Pinpointing Configuration Options}
	The synthetic setup has shown that our algorithm can conceptually handle both time and space. We were able to precisely identify commits for which performance changed substantially. By contrast, we were not able to fully reproduce this in a setting with real-world data regarding pinpointing change points to configuration options. Possible explanations in the case of \textsc{lrzip} include a partly inconsistent configuration interface. 
	The reason for that is not a limitation of our approach, but missing information in the real-world data (which makes the synthetic setup so important). So, we have documented that, given a precisely identified commit, we can link the affected files to configuration options with little effort. In turn, given a set of code segments of a commit, matching this with the overall code base has been extensively studied before under the umbrella of feature location techniques~\cite{hill_which_2013}.
	Inconsistencies in the documentation of configuration options is a well known and prevalent problem~\cite{rabkin_static_2011}. Our approach might be a means to complement existing feature location techniques with information about configuration-dependent performance changes. 	
	{\color{red} In our latter experiment, instead of directly sampling new arbitrary configurations, we used a finite set of configurations as a proxy (cf.~Section~\ref{sec:rq2_operationalization}). This does not reflect the original layout of our algorithm, but lets us efficiently explore some limitations of this approach. We have learned that for interpreting the results of our approach, domain knowledge is advantageous as we have identified incomplete and inconsistent configuration interfaces as possible limiting factors. We acknowledge that, while our approach performs well as an automated pipeline in a controlled environment, for a real-world applications, a semi-automated setup with additional manual measurement prioritization and deeper knowledge of configuration constraints would improve the applicability of our approach.} This is a promising further avenue of future work.

@@

	Regarding \emph{external validity}, we cannot claim that we have identified all limitations possible in a practical setting. However, we selected popular software systems that are well optimized for performance and often make use of configuration options to tune performance.  {\color{red}The results for the experiment with these software systems do no allow for generalization, but illustrate limitations faced in a practical setting. Since these results, in conjunction with our exhaustive synthetic study, are reproducible, we were able to test corner cases and assess scalability so that we believe that our results hold for many practical use cases.}

	\section{Related Work}\label{sec:rw}
	\paragraph{Performance-relevant Commits} Identifying performance-changing commits is a prevalent challenge in regression testing, often applying change point detection algorithms on batches of performance observations~\cite{cityIdentifying2014,daly_industry_2020}. A way to reduce testing overhead is to employ only a fraction of performance observations or prioritize commits. Sandoval Alcocer et al. assess performance for uniformly selected commits to estimate the risk of a performance change introduced by unobserved commits~\cite{sandoval_alcocer_learning_2016,alcocer_prioritizing_2020}. Huang et al. estimate the risk of performance changes solely based on static analysis of individual commits to prioritize commits for regression testing without assessing performance~\cite{huang_performance_2014}. Mühlbauer et al. use Gaussian Processes (GP) and iterative sampling to learn performance histories with few performance observations~\cite{muhlbauer_accurate_2019}. They extract performance changing commits by applying change point detection algorithms to learned performance histories. {\color{red}Nonetheless, they do not associate change points with individual configuration options or interactions.} All of the above approaches reduce testing overhead, but do not address the performance across different configurations of software systems. Our approach incorporates performance changes across different configurations by leveraging similarities in related configuration's performance histories.
	\paragraph{Performance of Configurable Software Systems} Associating configuration options and interactions with performance a conceptual basis for our work. There exists extensive work on modeling configuration options as features for machine learning, such as classification and regression trees~\cite{guoVariabilityawarePerformancePrediction2013,sarkarCostEfficientSamplingPerformance,nairUsingBadLearners2017}, multi-variable regression~\cite{siegmundPerformanceinfluenceModelsHighly2015}, and deep neural networks~\cite{haDeepPerf2019}. In contrast to learning performance for arbitrary configurations, our approach aims at explaining performance change, and subsequently, a change in a configuration option's or interaction's influence on performance over time. A similar scenario to ours has been studied by Jamshidi et al.~\cite{jamishidi_transfer_2017,jamshidi_transfer_gp_2017,jamshidi_learning_2018}. Between software versions, the performance influence of only few configuration options changes. The authors propose two approaches for learning a transfer function between performance models of different environments, such as versions, based on GPs~\cite{jamshidi_transfer_gp_2017} and progressive shrinking the configuration space~\cite{jamshidi_learning_2018}. We see great potential for future work in applying GPs to our problem statement as they have been successfully applied for the two orthogonal perspectives of time and configuration space.

>>>>

	Our main idea is as follows: We address the configuration complexity of this problem by selecting representative sample sets of configurations.
	Next, we uniformly sample a constant number of commits for each configuration and conduct respective performance measurements.
	Based on these measurement data, we assemble a prediction model.
	For each configuration, we estimate the likelihood of each commit being a change point, i.e. that performance of some configurations changes abruptly compared to the previous commit.
	Next, we leverage similarities in the performance histories of configurations that share common options.
	Since we sample the commits for each configuration independently, for each change point, we obtain many estimations using measurements of different commits.
	Overall, this allows us to obtain more accurate estimations of performance-changing commits with tractable effort.
	Based on a mapping from configurations to predicted change point probabilities for each configuration and commit, we derive the options responsible for each particular change point: configuration-dependent change points.
	In summary, we offer the following contributions:
	\begin{itemize}
		\item A novel technique to effectively identify shifts in performance of configurable software systems. We are able to pinpoint causative commits and affected configuration options with high accuracy and tractable effort.
		\item A feasibility demonstration of our approach by implementing an adaptive learning strategy to obtain accurate estimations with acceptable measurement cost.
		\item An evaluation using both synthetic and real-world performance data from three configurable software systems. Synthetic data lets us assess our model and approach conceptually and at scale, whereas we are able to assess practicality with real-world data.
		\item A companion website\footnote{\url{https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions} or archived at {\color{red}\url{https://archive.softwareheritage.org/browse/origin/https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions}}} providing supplementary material including a reference implementation of our approach, performance measurement data, and additional visualizations.
	\end{itemize}

@@

	\begin{figure}
		\includegraphics[width=0.5\textwidth]{images/sampling_plan}

		\caption{Active sampling strategies for a DBMS with 4 configurations}
		\label{fig:sampling plan}
	\end{figure}

@@

	To obtain accurate estimations with a limited budget of measurements, we require a different approach. We propose an iterative and adaptive sampling approach. That is, we use an initial, but small sample set to explore the problem space and then increase the level of granularity at promising regions and dimensions (development history segments as well as individual features and interactions, in our case). 
	Considering introductory DBMS example (cf. Figure~\ref{fig:dbms_performance}), we illustrate our sampling strategy in Figure~\ref{fig:sampling plan}. Here, each box depicts a possible measurements, i.e., pair of configuration and commit. The black bars represent change points hidden in the software system‘s performance histories, i.e., measurements immediately before and after a change point are dissimilar. The current state of the sample set comprises all measurements that are filled in grey.

	To include new measurements into the sample set, we take into account four possible objectives. On the one hand, our sample set might already hint to some possible change points and associated configuration options. On the other hand, our sample set might be too sparse and we cannot infer some change points yet. To address this trade-off, we split our objectives into exploitation and exploration.

	 For \emph{exploitation}, we sample new measurements to verify guesses based on the current state. That is, in Figure~\ref{fig:sampling plan}, we might include the measurements depicted by black-filled symbols. The black-filled stars represent measurements are of interest because we have already identified one change point for the configuration \textsc{Compression}. We exploit this knowledge to test for further configuration options involved in this change point. The black-filled pentagons represent measurements of interest, because we have identified another early change point for \textsc{Compression}, but within a rather broad range of commits. Thus, we include further measurements from that range to narrow down the possible range for this particular change point.

	 For \emph{exploration}, we might include the measurements by non-filled symbols. The stars represent measurements that explore new configurations. In our example, the interaction between \textsc{Compression} and \textsc{Encryption} is the only configuration left to be measured. We include several commits of a particular new configuration to unveil possible performance variation that indicate possible change points. The pentagons represent measurements that are of interest because they increase the overall measurement coverage. We include measurements from large intervals of not yet measured commits since possible change points can be hidden here.

	These four strategies (exploration and exploitation for configurations and commits) prioritize measurements to be included nex into the sample set. Adaptive sampling techniques have been successfully applied to obtain both performance-influence models~\cite{siegmundPredictingPerformanceAutomated2012,sarkarCostEfficientSamplingPerformance} and performance histories\cite{muhlbauer_accurate_2019} before. However, it is unclear whether fewer measurements are sufficient to assess the performance-influence of configuration options and interactions with respect to version changes. This is what we aim to accomplish with our work.

	\begin{comment}
	{\color{magenta}We illustrate the measurement effort in Figure~\ref{fig:sampling_all}, where each dot represents a pair of configuration and version.
		To reduce measurement effort, one could limit the measurements to either a few configurations (cf. Figure~\ref{fig:sampling_configurations}) or a few commits (cf. Figure~\ref{fig:sampling_versions}), but at the cost of information loss along the respective complementary dimension.
		This would leave us with either accurate performance-influence models for few versions, resulting in poor temporal granularity, or with accurate performance histories, but only for a few configurations.
		Both do not provide accurate estimations of changes in the performance influence.}

	To obtain accurate estimations with a limited budget of measurements, we require a different approach. 
	The latter describes, in fact, an iterative and adaptive sampling approach.
	That is, we use an initial, but sparse sample set to explore the problem space and then increase the level of granularity at promising regions and dimensions (history segments and individual features and interactions, in our case).
	Adaptive sampling techniques have been successfully applied to  obtain both performance-influence models~\cite{siegmundPredictingPerformanceAutomated2012,sarkarCostEfficientSamplingPerformance} and performance histories~\cite{muhlbauer_accurate_2019} before.
	However, it is unclear whether fewer measurements are sufficient to assess the performance-influence of configuration options and interactions with respect to version changes. This is what we aim to accomplish with our work.
	\end{comment}

	\begin{comment}
	\begin{figure}
		\captionsetup{format=hang}
		\begin{subfigure}{0.49\linewidth}
			\includegraphics[width=\linewidth]{images/sampling_1.pdf}
			\caption{Many configurations, but few versions}
			\label{fig:sampling_configurations}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.49\linewidth}
			\includegraphics[width=\linewidth]{images/sampling_2.pdf}
			\caption{Many versions, but few configurations}
			\label{fig:sampling_versions}
		\end{subfigure}
		\vfill
		\begin{subfigure}{0.49\linewidth}
			\includegraphics[width=\linewidth]{images/sampling_3.pdf}
			\caption{Many versions and configurations\\}
			\label{fig:sampling_all}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.49\linewidth}
			\includegraphics[width=\linewidth]{images/sampling_4.pdf}
			\caption{Initial sample augmented with additional measurements}
			\label{fig:sampling_both}
		\end{subfigure}

@@

	%	\begin{figure}
	%\centering
	%\includegraphics[width=\linewidth]{images/complexity.pdf}
	%\caption{Complexity}
	%\label{fig:complexity}
	%\end{figure}
	\end{comment}

@@

	\subsubsection{Candidate Commits}
	For each configuration, we compute an approximation of the change point likelihood over commits (cf. Figure~\ref{fig:cpp_example}) to identify local maxima (peaks). We select such peaks under the condition that the peak change point likelihood is greater than a  threshold $t_\text{CPL}$, as defined in Equation~\ref{eq:t_cpl}. This threshold is the average change point likelihood over all commits plus a factor $N_\text{CPL}$ times its corresponding standard deviation over all commits. 
	\begin{equation}
	\begin{gathered}
	t_\text{CPL} = \frac{1}{\vert V\vert} + N_\text{CPL} \cdot \sqrt{\frac{1}{\vert V\vert} \cdot \sum_{v \in V} \Big(\frac{1}{\vert V\vert} - p'(v)\Big)^2}
	\label{eq:t_cpl}
	\end{gathered}
	\end{equation}
	The factor $N_\text{CPL}$ (default value: 3) allows us to filter peaks that do not stand out enough and might be false positives. The greater $N_\text{CPL}$ is, the stricter the filtering of peaks. That is, for each configuration in our learning set, we obtain a set of commits (configuration-specific candidate commits) that represent possible change points.
	To reduce variation among the obtained commits, we cluster the commits using kernel density estimation (KDE). The purpose of KDE in our setting is to estimate the probability mass function of whether a commit is a change point. The local maxima of this KDE, subsequently, represent our set of candidate commits. Note that if two distinct change points are almost coincident commit-wise, they can be mistakenly identified by our approach as one change point that, subsequently, can be associated with configuration options belonging to the two individual change points.

@@

	\subsubsection{Acquiring Configurations}
	The exploration of new configurations is similar to the initial sample selection strategy (distance-based sampling) described in Section~\ref{sec:initialization}. We select a constant number of configurations (default value: 75) for exploration.

	The exploitation part of acquiring new configurations is guided by the current set of candidate solutions. Each candidate solution describes a change point and associated configuration options. These associated options can be a correct assignment or be an over- or under-approximation. In the latter cases, too many or too few options are associated with a change point. We exploit the existing candidate solutions in a way that addresses both under- as well as over-approximation. We select new configurations using a constraint solver and therefore can specify additional constraints. 
	In the case of under-approximation, too few relevant configuration options are associated with a change point. Given a candidate change point, we require that all associated options keep enabled and that 50\,\% of all not-associated options can be selected. In doing so, we keep associated options, but allow new configuration options to be included. Likewise, in the case of over-approximation, to many irrelevant configuration options are associated with a change point. Given a candidate change point, we require that variation only occurs among 50\,\% of the options associated with the change point. That is, we can remove up to 50\,\% of configuration options and narrow down the selection of relevant configuration options.

@@

	To mitigate this trade-off in our evaluation, we conduct two separate experiments, based on synthetic and real-world performance measurements. The first experiments uses synthesized performance data and allows for a controlled experimental setup to assess scalability, accuracy, and efficiency at low cost while simultaneously be able to simulate different scenarios by varying the number of change points and affected configurations in the synthesized data.
	The second experiment uses a batch of real-world performance measurements of three software systems as a non-exhaustive ground truth to explore whether our algorithm can be practically applied to real-world systems.

@@

	As a workload for the file compression tools \textsc{lrzip} and \textsc{xz}, we compressed the Silesia corpus which contains over 200~MB of files of different types.  This workload was designed to compare different file compression algorithms and tools. For the audio transcoder \textsc{oggenc}, we encoded a raw WAVE audio file of over 60~MB from the Wikimedia Commons collection.  The workloads for \textsc{xz} and \textsc{lrzip} have been studied in previous work~\cite{siegmund_views_2015,muhlbauer_accurate_2019}. For all three subject systems, we assess performance by reporting the execution time.

@@

	\subsubsection{Operationalization}\label{sec:rq2_operationalization}
	For the experiment, we apply our approach on the performance measurements obtained in our pre-study with a few adjustments.
	First, although our study provides a broad and representative sample set of configurations, we cannot arbitrarily sample valid configurations for a couple of reasons. In particular, these include the commits that do not build as well as hidden configuration options and constraints. While we collected configuration options and constraints thoroughly from documentation artifacts, we can not assure that our selection is complete. We discuss this limitation further in the threats to validity.

	Instead, when acquiring new configurations (cf. Sec.~\ref{sec:acquisition}), we select configurations from our batch of measurements that (1) have not already been used by our algorithm, and (2) are "closest" to the requested configuration (i.e., with the minimal Hamming distance). The rationale is that this allows for rapid repetitions with different initializations of our algorithm. Since we employ multiple sampling strategies in our pre-study we are confident that our broad batch of measurements is representative. Second, we set the number of measurements per iteration to 300. This value is substantially smaller than in our controlled experiment setup, where we explore the scalability of our approach. Third, as an initial configuration sample set, we randomly sample 5 configurations from our pre-study set. 
	Last, we initialize our approach with a relative performance threshold $\theta_\tau$ (cf. Equation~\ref{eq:theta}) of ten percent. In the context of the reported relative variation among performance measurements on the machines used, we consider this a rather conservative threshold. 
	We repeat the experiments 10 times with different seeds to assess robustness of our approach and account for the randomness in exploring the configuration space.
	We assess practicality by reporting measurement effort, number of iterations in relation to the pre-study's results, which are based on a vastly greater measurement budget.

@@

	\section{Discussion}
	\paragraph{Efficiency} Among all experiments conducted with our synthetic setup, our approach yielded reasonable to high accuracy across most parameter settings. A subsequent analysis of parameter settings showed that our algorithm is not limited by the size of a problem space, but rather the number of measurements per iteration. Our algorithm found multiple change points across settings of up to 2,500 commits and and 2$^{64}$ configurations, resulting in a search space of about $4.6 \cdot 10^{22}$ possible measurements.
	In the context of not more than 5,000 measurements per iteration, the maximum number of measurements (after 30 iterations) added to the initial set of measurements is 150,000. For a setting with 2,500 commits, this represents around $1.4 \cdot 10^{-6}$\,\% and $3.2\cdot 10^{-16}$\,\% of possible measurements for settings with 32 and 64 options, respectively. The vast majority of cases terminated in early iterations. Hence, we are confident that our approach scales and can efficiently approximate change points for large software systems. We, we are able to reproduce this finding with real-world performance when identifying the temporal location of change points. For our controlled experiment setup, we have not tuned the algorithms’ parameters to reach optimal efficiency or accuracy. However, even without tuning, we demonstrated the feasibility and efficiency of our approach. Project-specific fine tuning of several parameters is a possible extension to our approach and can only further improve our results.

	\paragraph{Pinpointing Configuration Options}
	The synthetic setup has shown that our algorithm can conceptually handle both time and space. We were able to precisely identify commits for which performance changed substantially. By contrast, we were not able to fully reproduce this in a setting with real-world data regarding pinpointing change points to configuration options. Possible explanations in the case of \textsc{lrzip} include a partly inconsistent configuration interface. 
	The reason for that is not a limitation of our approach, but missing information in the real-world data (which makes the synthetic setup so important). So, we have documented that, given a precisely identified commit, we can link the affected files to configuration options with little effort. In turn, given a set of code segments of a commit, matching this with the overall code base has been extensively studied before under the umbrella of feature location techniques~\cite{hill_which_2013}.
	Inconsistencies in the documentation of configuration options is a well known and prevalent problem~\cite{rabkin_static_2011}. Our approach might be a means to complement existing feature location techniques with information about configuration-dependent performance changes. 	
	In our latter experiment, instead of directly sampling new arbitrary configurations, we used a finite set of configurations as a proxy (cf.~Section~\ref{sec:rq2_operationalization}). This does not reflect the original layout of our algorithm, but lets us efficiently explore some limitations of this approach. We have learned that for interpreting the results of our approach, domain knowledge is advantageous as we have identified incomplete and inconsistent configuration interfaces as possible limiting factors. We acknowledge that, while our approach performs well as an automated pipeline in a controlled environment, for real-world applications, a semi-automated setup with additional manual measurement prioritization and deeper knowledge of configuration constraints would improve the applicability of our approach. This is a promising further avenue of future work.

@@

	Regarding \emph{external validity}, we cannot claim that we have identified all limitations possible in a practical setting. However, we selected popular software systems that are well optimized for performance and often make use of configuration options to tune performance.  Although we have shown possible limitations in a practical setting, our results, in conjunction with our exhaustive synthetic study, are reproducible, and we were able to test corner cases and assess scalability so that we believe that our results hold for many practical use cases.

	\section{Related Work}\label{sec:rw}
	\paragraph{Performance-relevant Commits} Identifying performance-changing commits is a prevalent challenge in regression testing, often applying change point detection algorithms on batches of performance observations~\cite{cityIdentifying2014,daly_industry_2020}. A way to reduce testing overhead is to employ only a fraction of performance observations or prioritize commits. Sandoval Alcocer et al. assess performance for uniformly selected commits to estimate the risk of a performance change introduced by unobserved commits~\cite{sandoval_alcocer_learning_2016,alcocer_prioritizing_2020}. Huang et al. estimate the risk of performance changes solely based on static analysis of individual commits to prioritize commits for regression testing without assessing performance~\cite{huang_performance_2014}. Mühlbauer et al. use Gaussian Processes (GP) and iterative sampling to learn performance histories with few performance observations~\cite{muhlbauer_accurate_2019}. They extract performance changing commits by applying change point detection algorithms to learned performance histories. Nonetheless, they do not associate change points with individual configuration options or interactions. All of the above approaches reduce testing overhead, but do not address the performance across different configurations of software systems. Our approach incorporates performance changes across different configurations by leveraging similarities in related configuration's performance histories.
	\paragraph{Performance of Configurable Software Systems} Associating configuration options and interactions with performance a conceptual basis for our work. There exists extensive work on modeling configuration options as features for machine learning, such as classification and regression trees~\cite{guoVariabilityawarePerformancePrediction2013,sarkarCostEfficientSamplingPerformance,nairUsingBadLearners2017}, multi-variable regression~\cite{siegmundPerformanceinfluenceModelsHighly2015}, and deep neural networks~\cite{haDeepPerf2019}. In contrast to learning performance for arbitrary configurations, our approach aims at explaining performance change, and subsequently, a change in a configuration option's or interaction's influence on performance over time. A similar scenario to ours has been studied by Jamshidi et al.~\cite{jamishidi_transfer_2017,jamshidi_transfer_gp_2017,jamshidi_learning_2018}. Between software versions, the performance influence of only few configuration options changes. The authors propose two approaches for learning a transfer function between performance models of different environments, such as versions, based on GPs~\cite{jamshidi_transfer_gp_2017} and progressive shrinking the configuration space~\cite{jamshidi_learning_2018}. We see great potential for future work in applying GPs to our problem statement as they have been successfully applied for the two orthogonal perspectives of time and configuration space.
